### 2022-02-15 manuscript/main.tex
Updates to main manuscript, re-formatted PDF located in 'wiley' subdirectory to reflect QJRMS preprint formatting.

% Comment: try to couple the local and the mesoscale here
The meteorology of extreme heat events and its impacts on urban areas can be observed from the synoptic and local scales. From a synoptic scale, extreme heat events are often caused by the sustained presence of a high-pressure system over an area, resulting in lower wind speeds and warm air subsidence, promoting higher surface temperatures \citep{black2004, miralles2014}. From a local perspective, the amplified impact of extreme heat events on cities is a result of the urban heat island (UHI) effect, which occurs as a result of the modification of land surface properties due to the built environment. The modification of surface properties has been shown to increase near-surface air temperatures due to factors such as radiation entrapment, increased heat storage, and lower evapotranspirative cooling \citep{chen2014, li2013, ramamurthy2017, zhao2018}. Additionally, urban areas near large bodies of water experience effects from the sea breeze, which has been shown to play a moderating influence on the intensity of the UHI effect \citep{hu2016, jiang2019, stefanon2014}. The processes on these two scales can be connected by understanding the structure and dynamics of the urban boundary layer (UBL), which is the lowest part of the troposphere in which surface-atmosphere exchanges occur that directly affect human activity. There have been a large number of numerical studies performed to improve our understanding of UBL processes during extreme heat events, which have been important for conceptualizing the role of synoptic-scale and surface forcings on urban climate. Something about the vertical structure of the UBL and coupling these

@@

Observations of the UBL during extreme heat events are even more limited. \citet{ramamurthy2017} used microwave radiometers to observe the UBL over New York City in July 2016 to find that the UHI effect was amplified during heat wave events and that spatial variability throughout the city was significant throughout the observation period. \citet{jiang2019} explores the effects of heat waves on rural and urban areas for several cities in China using ground-based observations with a focus on the UHI effect, finding that the effect was amplified during heat waves due to greater surface solar radiation and shifts in wind direction contributing to advection of heated air masses over the studied cities. \citep{wu2019} uses a combination of a ceilometer and multiple lidars to observe the evolution of UBL structure, air quality, and pollutant transport during a heat wave in New York City, demonstrating sharp rates of UBL growth due to convective activity and an increase of pollutant concentration and regional transport. \citet{zhang2020} uses aircraft-based observations to provide a comprehensive analysis of UBL structure during heat wave events over cities in the United States throughout a 10-year period, providing insights into the 'heat dome' thermodynamic structure over cities and the variability between heat wave events due to local (such as surface properties in urban areas) and large-scale (such as synoptic meteorological conditions) forcings.

@@

\subsubsection{Horizontal winds}
Extreme heat events coincided with an overall reduction of horizontal wind speeds ($U$) in the UBL. More specifically, the magnitude of $U$ during extreme heat events is similar in magnitude to $U$ during normal days with the exception of early morning hours and at upper levels of the UBL. As shown in Figure \ref{fig:extreme-heat-normal-comparison-contours-U}, modest reductions in $U$ ($-1.2 \leq \sigma \leq -0.4$) during extreme heat events are present throughout the UBL from early to mid-morning, with little differences throughout the rest of the day ($-0.4 \leq \sigma \leq 0.4$). Larger deviations between $U$ values are present at the top of the UBL where synoptic conditions become dominant.
\\ \\
Vertical profiles of $U$ for normal and extreme heat events provide a more detailed view of the differences in UBL structure. Across all sites, $U$ is similar throughout the UBL overnight, afternoon, and evenings. During early morning hours, extreme heat event $U$ values decrease by 25 to 50\% throughout the entire UBL (see Figures \ref{fig:vertical_profiles-heat_wave-normal-U-bron}, \ref{fig:vertical_profiles-heat_wave-normal-U-quee}, \ref{fig:vertical_profiles-heat_wave-normal-U-stat}), although both event types present a classical logarithmic wind profile, with surface friction effects present through \SI{500}{\meter}. Another phenomenon worth noting is the difference in $U$ profiles above \SI{2000}{\meter}; profiles of $U$ during extreme heat events are more consistent between sites and vertically than during normal days. This phenomenon may highlight the effect of synoptic meteorological conditions on $U$, as the UBL typically remains below \SI{2500}{\meter}. During extreme heat events, anticyclonic conditions produce more stable atmospheric conditions relative to normal days, resulting in less variability between heat events than during normal days.
\\ \\
Extreme heat events result in a southwesterly shift in $U$ throughout the UBL. This shift is present most evidently closer to the surface, as shown in Figures \ref{fig:wind_rose-bron}, \ref{fig:wind_rose-quee}, and \ref{fig:wind_rose-stat}, with winds at \SI{100}{\meter} coming primarily from the southwest quadrant. Figure \ref{fig:wind_rose-quee} shows that Queens also presents a secondary maximum with winds approaching from the south, which suggests effects from the Atlantic sea breeze (effects from the sea breeze will be further discussed in Section \ref{section:sea_breeze_effects}). At \SI{1000}{\meter}, the directionality of prevailing winds becomes more uniform between normal and extreme heat days, as winds primarily approach New York City from the west-southwest. The disparity in wind directions between 100 and \SI{1000}{\meter} suggests that localized wind fields play a major role in UBL dynamics at lower levels whereas synoptic-scale atmospheric conditions increasingly dominate with increasing height.

\subsubsection{Vertical motion}
On average, extreme heat events do not appear to produce significant changes in vertical velocity ($w$). Figure \ref{fig:extreme-heat-normal-comparison-100m-w} shows average diurnal profiles of $w$ at all locations at \SI{100}{\meter} above ground level, with similar mean values throughout the day between normal days and extreme heat events. During extreme heat events, however, the variability of $w$ is less in the early morning hours and greater in the evening. This phenomenon is also observed in vertical profiles of $w$ at all locations as shown in Figures \ref{fig:extreme-heat-normal-vertical_profiles-w-bron},\ref{fig:extreme-heat-normal-vertical_profiles-w-quee}, and \ref{fig:extreme-heat-normal-vertical_profiles-w-stat}. At all locations, overnight and morning profiles of $w$ (0:00 and 6:00 LST) show significantly lower variability in $w$ throughout the UBL with similar magnitudes of mean $w$. Despite similar means and deviations in the early afternoon (12:00 LST), evening profiles (18:00 LST) show significantly higher variability in $w$ below \SI{500}{\meter} for all sites, with the Bronx showing this occurrence extend through the UBL.

%%%%% Discuss the observed sea breeze circulation
\section{Effects of the sea breeze circulation} \label{section:sea_breeze_effects} 
Sea breezes in New York City occur as a result of land-sea temperature gradients from two arms of the Atlantic Ocean; the New York Bight to the southeast and Long Island Sound to the northeast. Sea breezes from both bodies increase the complexity of UBL dynamics over New York City due to the coalescence of opposing fronts over a complex topography \citep{bornstein1981}. A typical sea breeze event in New York City is defined by calm ambient low-level winds ($\leq$ \SI{5}{\meter\per\second}, the formation of a large land-sea temperature gradient in the mid- to late morning, strong late-morning thermals that promote low-level convergence, and afternoon to early-evening onshore moisture transport and reduction in surface air temperatures (especially in areas closest to the shore) \citep{childs2005, frizzola1963, gedzelman2003}. 
\\
Sea breeze events occurred on approximately 56\% of all days observed. The high frequency of occurrence is likely attributable to the large land-sea temperature gradient that is common during warmer months \citep{gedzelman2003}, as days were chosen exclusively between May and September. Maximum land-sea surface temperature differences during days with identifiable sea breeze events averaged at \SI{12}{\kelvin}, with a strong diurnal profile with the peak difference occurring around midday (see Figure \ref{fig:sea_breeze-lst_sst-diff}). The frequency of occurrence increases when observing days during extreme heat events, as the lack of a strong synoptic wind allows for the sea breeze circulation to become dominant in the metropolitan area \citep{miller2003}.

@@

\subsection{UBL dynamics during sea breeze events}
Days with identifiable sea breeze events had lower $U$ throughout the majority of the UBL, with the most significant decreases during the nighttime, potentially due to the lessening of onshore flow due to the reduction of the land-sea temperature gradient \citep{pullen2007}, as shown in Figure \ref{fig:sea_breeze-lst_sst-diff}. Vertical motions, however, increased significantly in the Bronx and Queens during the late morning and early afternoon, as shown in Figure \ref{fig:sea_breeze_heat_wave_anomaly-w}. These anomalies likely indicate the increased presence of updrafts in urbanized areas.

During days with identified sea breeze circulations, easterly winds increase in frequency in the lower levels of the UBL, as shown in Figures \ref{fig:wind_direction-heat_wave-sea_breeze-histogram-bron}, \ref{fig:wind_direction-heat_wave-sea_breeze-histogram-quee}, and \ref{fig:wind_direction-heat_wave-sea_breeze-histogram-stat}. Southeasterly winds (due to sea breezes from the New York Bight) increased in frequency compared to all other days at all locations. The occurrence frequency of southeasterly winds is correlated with the distance between the observation site and the largest body of water in proximity of the metropolitan area (Atlantic Ocean), as Staten Island reported 92.1\% of all winds at \SI{100}{\meter} as southeasterly between 12:00 and 20:00 LST (distance of \SI{6.50}{\kilo\meter} from Lower New York Bay), whereas Queens reported 67.4\% (distance of \SI{16.5}{\kilo\meter}), and Bronx reported 55.6\% (distance of \SI{32.9}{\kilo\meter}).

For sites near Long Island Sound (the Bronx and Queens), northeasterly winds increased in frequency as well, though not to the same magnitude as southeasterly winds. This disparity in magnitude suggests that the Long Island Sound sea breeze front is weaker than the New York Bight sea breeze front, which aligns with previous studies of sea breeze fronts over New York City \citep{frizzola1963, meir2013}.

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\textwidth]{figs/heat_wave-normal-U-anomaly.png}
	\caption{Anomalies of $U$ during extreme heat events relative to the climatology over the urban boundary layer.}
	\label{fig:extreme-heat-normal-comparison-contours-U}
\end{figure}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\textwidth]{figs/heat_wave-normal-U-anomaly-100m.png}
	\caption{Anomalies of low-level $U$ during extreme heat events relative to the climatology.}
	\label{fig:extreme-heat-normal-100m-comparison-U}
\end{figure}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\textwidth]{figs/vertical_profiles-heat_wave-normal-U-bron.png}
	\caption{Vertical profiles of $U$ in the Bronx during normal days (blue) and extreme heat events (red).}
	\label{fig:vertical_profiles-heat_wave-normal-U-bron}
\end{figure}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\textwidth]{figs/vertical_profiles-heat_wave-normal-U-quee.png}
	\caption{Vertical profiles of $U$ in Queens during normal days (blue) and extreme heat events (red).}
	\label{fig:vertical_profiles-heat_wave-normal-U-quee}
\end{figure}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\textwidth]{figs/vertical_profiles-heat_wave-normal-U-stat.png}
	\caption{Vertical profiles of $U$ in Staten Island during normal days (blue) and extreme heat events (red).}
	\label{fig:vertical_profiles-heat_wave-normal-U-stat}
\end{figure}

>>>>

% Comment: try to couple the local and the mesoscale here
The meteorology of extreme heat events and its impacts on urban areas can be observed from the synoptic and local scales. From a synoptic scale, extreme heat events are often caused by the sustained presence of a high-pressure system over an area, resulting in lower wind speeds and warm air subsidence, promoting higher surface temperatures \citep{black2004, miralles2014}. From a local perspective, the amplified impact of extreme heat events on cities is a result of the urban heat island (UHI) effect, which occurs as a result of the modification of land surface properties due to the built environment. The modification of surface properties has been shown to increase near-surface air temperatures due to factors such as radiation entrapment, increased heat storage, and lower evapotranspirative cooling \citep{chen2014, li2013, ramamurthy2017a, zhao2018}. Additionally, urban areas near large bodies of water experience effects from the sea breeze, which has been shown to play a moderating influence on the intensity of the UHI effect \citep{hu2016, jiang2019, stefanon2014}. The processes on these two scales can be connected by understanding the structure and dynamics of the urban boundary layer (UBL), which is the lowest part of the troposphere in which surface-atmosphere exchanges occur that directly affect human activity. There have been a large number of numerical studies performed to improve our understanding of UBL processes during extreme heat events, which have been important for conceptualizing the role of synoptic-scale and surface forcings on urban climate. Something about the vertical structure of the UBL and coupling these

@@

Observations of the UBL during extreme heat events are even more limited. \citet{ramamurthy2017a} used microwave radiometers to observe the UBL over New York City in July 2016 to find that the UHI effect was amplified during heat wave events and that spatial variability throughout the city was significant throughout the observation period. \citet{jiang2019} explores the effects of heat waves on rural and urban areas for several cities in China using ground-based observations with a focus on the UHI effect, finding that the effect was amplified during heat waves due to greater surface solar radiation and shifts in wind direction contributing to advection of heated air masses over the studied cities. \citep{wu2019} uses a combination of a ceilometer and multiple lidars to observe the evolution of UBL structure, air quality, and pollutant transport during a heat wave in New York City, demonstrating sharp rates of UBL growth due to convective activity and an increase of pollutant concentration and regional transport. \citet{zhang2020} uses aircraft-based observations to provide a comprehensive analysis of UBL structure during heat wave events over cities in the United States throughout a 10-year period, providing insights into the 'heat dome' thermodynamic structure over cities and the variability between heat wave events due to local (such as surface properties in urban areas) and large-scale (such as synoptic meteorological conditions) forcings.

@@

\subsubsection{Horizontal winds}
Extreme heat events coincided with an overall reduction of horizontal wind speeds ($U$) in the UBL (Figure \ref{fig:extreme-heat-normal-100m-comparison-U}. More specifically, the magnitude of $U$ during extreme heat events is similar in magnitude to $U$ during normal days with the exception of early morning hours and at upper levels of the UBL. As shown in Figure \ref{fig:extreme-heat-normal-comparison-contours-U}, modest reductions in $U$ ($-1.2 \leq \sigma \leq -0.4$) during extreme heat events are present throughout the UBL from early to mid-morning, with little difference throughout the rest of the day ($-0.4 \leq \sigma \leq 0.4$). Larger deviations between $U$ values are present at the top of the UBL where synoptic conditions become dominant.
\\ \\
Vertical profiles of $U$ for normal and extreme heat events provide a more detailed view of the differences in UBL structure. Across all sites, $U$ is similar throughout the UBL overnight, afternoon, and evenings. During early morning hours, however, extreme heat event $U$ values decrease by 25 to 50\% throughout the entire UBL (see Figures \ref{fig:vertical_profiles-heat_wave-normal-U-bron}, \ref{fig:vertical_profiles-heat_wave-normal-U-quee}, \ref{fig:vertical_profiles-heat_wave-normal-U-stat}), although both event types present a classical logarithmic wind profile, with surface friction effects present through \SI{500}{\meter}. The reduction in $U$ during extreme heat events is likely due to the presence of an anticyclonic circulation that suppresses the nocturnal low-level jet over New York City \citep{chen1993}. Another phenomenon worth noting is the difference in $U$ profiles above \SI{2000}{\meter}; profiles of $U$ during extreme heat events are more consistent between sites and vertically than during normal days. This phenomenon is the effect of synoptic meteorological conditions on $U$, as the UBL typically remains below \SI{2500}{\meter}. During extreme heat events, anticyclonic conditions produce more consistent atmospheric conditions relative to normal days, resulting in less variability between heat events than during normal days.
\\ \\
Extreme heat events result in a southwesterly shift in $U$ throughout the UBL. This shift is present most evidently closer to the surface, as shown in Figures \ref{fig:wind_rose-bron}, \ref{fig:wind_rose-quee}, and \ref{fig:wind_rose-stat}, with winds at \SI{100}{\meter} coming primarily from the southwest quadrant. Figure \ref{fig:wind_rose-quee} shows that Queens also presents a secondary maximum with winds approaching from the south, which suggests effects from the Atlantic sea breeze (effects from the sea breeze will be further discussed in Section \ref{section:sea_breeze_effects}). At \SI{1000}{\meter}, the directionality of prevailing winds becomes more uniform between normal and extreme heat days, as winds primarily approach New York City from the west-southwest. The disparity in wind directions between 100 and \SI{1000}{\meter} suggests that localized wind fields play a more significant role in UBL dynamics at lower levels whereas synoptic-scale atmospheric conditions increasingly dominate with increasing height. Regardless, the uniformity of wind direction during extreme heat relative to normal days indicates that synoptic-scale effects can play a larger role at lower levels due to advection from the continent, especially with regards to thermal advection that leads to the transport of heated inland air masses over New York City \citep{ramamurthy2017b}.

\subsubsection{Vertical motion}
On average, extreme heat events do not appear to produce significant changes in vertical velocity ($w$). Figure \ref{fig:extreme-heat-normal-comparison-100m-w} shows average diurnal profiles of $w$ at all locations at \SI{100}{\meter} above ground level, with similar mean values throughout the day between normal days and extreme heat events. During extreme heat events, however, the variability of $w$ is less in the early morning hours and greater in the evening. This phenomenon is also observed in vertical profiles of $w$ at all locations as shown in Figures \ref{fig:extreme-heat-normal-vertical_profiles-w-bron}, \ref{fig:extreme-heat-normal-vertical_profiles-w-quee}, and \ref{fig:extreme-heat-normal-vertical_profiles-w-stat}. At all locations, overnight and morning profiles of $w$ (0:00 and 6:00 LST) show significantly lower variability in $w$ throughout the UBL with similar magnitudes of mean $w$. Despite similar means and deviations in the early afternoon (12:00 LST), evening profiles (18:00 LST) show significantly higher variability in $w$ below \SI{500}{\meter} for all sites, with the Bronx showing this occurrence extend through the UBL. The similarity in vertical profiles of $w$ may be a result of a balance between large-scale subsidence (due to the presence of high-pressure during extreme heat events) and the effects of increased surface forcings during extreme heat events relative to normal days \citep{dong2018, zhang2009}.

%%% Prathap to add text regarding vertical motion, this area feels lacking

%%%%% Discuss the observed sea breeze circulation
\section{Effects of the sea breeze circulation} \label{section:sea_breeze_effects} 
Sea breezes in New York City occur as a result of land-sea temperature gradients from two arms of the Atlantic Ocean; the New York Bight to the southeast and Long Island Sound to the northeast. Sea breezes from both bodies increase the complexity of UBL dynamics over New York City due to the coalescence of opposing fronts over a complex topography \citep{bornstein1981}. A typical sea breeze event in New York City is defined by calm ambient low-level winds ($\leq$ \SI{5}{\meter\per\second}, the formation of a large land-sea temperature gradient in the mid- to late morning, strong late-morning thermals that promote low-level convergence, and afternoon to early-evening onshore moisture transport and reduction in surface air temperatures (especially in areas closest to the shore) \citep{childs2005, frizzola1963, gedzelman2003}. 
\\ \\
Sea breeze events occurred on approximately 56\% of all days observed. The high frequency of occurrence is  attributable to low-level convergence due to  the large land-sea temperature gradient that is common during warmer months \citep{childs2005, gedzelman2003, thompson2007}, as days were chosen exclusively between May and September. Maximum land-sea surface temperature differences during days with identifiable sea breeze events averaged at \SI{12}{\kelvin}, with a strong diurnal profile with the peak difference occurring around midday (see Figure \ref{fig:sea_breeze-lst_sst-diff}). The frequency of occurrence increases when observing days during extreme heat events, as the lack of a strong synoptic wind allows for the sea breeze circulation to become dominant in the metropolitan area \citep{miller2003}.

@@

\subsection{UBL dynamics during sea breeze events}
Days with identifiable sea breeze events had lower $U$ throughout the majority of the UBL, with the most significant decreases during the nighttime, potentially due to the lessening of onshore flow due to the reduction of the land-sea temperature gradient \citep{pullen2007}, as shown in Figure \ref{fig:sea_breeze-lst_sst-diff}. Vertical motions, however, increased significantly in the Bronx and Queens during the late morning and early afternoon, as shown in Figure \ref{fig:sea_breeze_heat_wave_anomaly-w}. These anomalies likely indicate the increased presence of updrafts in urbanized areas.
\\ \\
During days with identified sea breeze circulations, easterly winds increase in frequency in the lower levels of the UBL, as shown in Figures \ref{fig:wind_direction-heat_wave-sea_breeze-histogram-bron}, \ref{fig:wind_direction-heat_wave-sea_breeze-histogram-quee}, and \ref{fig:wind_direction-heat_wave-sea_breeze-histogram-stat}. Southeasterly winds (due to sea breezes from the New York Bight) increased in frequency compared to all other days at all locations. The occurrence frequency of southeasterly winds is correlated with the distance between the observation site and the largest body of water in proximity of the metropolitan area (Atlantic Ocean), as Staten Island reported 92.1\% of all winds at \SI{100}{\meter} as southeasterly between 12:00 and 20:00 LST (distance of \SI{6.50}{\kilo\meter} from Lower New York Bay), whereas Queens reported 67.4\% (distance of \SI{16.5}{\kilo\meter}), and Bronx reported 55.6\% (distance of \SI{32.9}{\kilo\meter}). 
\\ \\
For sites near Long Island Sound (the Bronx and Queens), northeasterly winds increased in frequency as well, though not to the same magnitude as southeasterly winds. This disparity in magnitude suggests that the Long Island Sound sea breeze front is weaker than the New York Bight sea breeze front, which aligns with previous studies of sea breeze fronts over New York City \citep{frizzola1963, meir2013}.

@@

\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\textwidth]{figs/heat_wave-normal-U-anomaly-100m.png}
	\caption{Anomalies of low-level $U$ during extreme heat events relative to the climatology.}
	\label{fig:extreme-heat-normal-100m-comparison-U}
\end{figure}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\textwidth]{figs/heat_wave-normal-U-anomaly.png}
	\caption{Anomalies of $U$ during extreme heat events relative to the climatology over the urban boundary layer.}
	\label{fig:extreme-heat-normal-comparison-contours-U}
\end{figure}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\textwidth]{figs/vertical_profiles-heat_wave-normal-U-bron.png}
	\caption{Vertical profiles of $U$ in the Bronx during normal days (blue) and extreme heat events (red).}
	\label{fig:vertical_profiles-heat_wave-normal-U-bron}
\end{figure}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\textwidth]{figs/vertical_profiles-heat_wave-normal-U-quee.png}
	\caption{Vertical profiles of $U$ in Queens during normal days (blue) and extreme heat events (red).}
	\label{fig:vertical_profiles-heat_wave-normal-U-quee}
\end{figure}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.75\textwidth]{figs/vertical_profiles-heat_wave-normal-U-stat.png}
	\caption{Vertical profiles of $U$ in Staten Island during normal days (blue) and extreme heat events (red).}
	\label{fig:vertical_profiles-heat_wave-normal-U-stat}
\end{figure}
